### 2020-01-22 DataWrangling/Images extraction.Rmd
Jan22.2020

# Extract images

```{r Images to Extract}
# write.csv(NoDelft[, c(21:26)], "NoDelft.csv")
# length(unique(NoDelft$TID_OMRS))
# write.csv(MDF[, c(21:26)], "MDF.csv")
# rm(NoDelft)

######### Extraction 1000 ######
# rm(Adult_Valid, CAD3_delft,CAD6_delft, DiffNotReceived, DiffReceived, Multi_TID_60K, Multi_TID_Master, n_occur, NotReceived, QA.BGD, ReceivedValid, ReceivedValid, Master_TID_Received, ExtractLuan, Extract)
# glimpse(Master_TID_Received)

# table(Master_TID_Received$Result.Date)

ExtractBGD1000 <- MDF[order(MDF$Result.Date),]
ExtractBGD1000 <- ExtractBGD1000[!duplicated(ExtractBGD1000$TID_OMRS), ]
ExtractBGD1000 <- subset(ExtractBGD1000, !is.na(ExtractBGD1000$Xpert2Outcome_num))

summary(ExtractBGD1000$Xpert2Outcome_num)

ExtractBGD1000 <- ExtractBGD1000[30263:29264, ]
length(unique(ExtractBGD1000$TID_OMRS))

names(ExtractBGD1000)
write.csv(ExtractBGD1000[, c(1,2,23)], "ExtractBGD1000.csv")

# BGD1000 <- read_csv("BGD1000.csv")
glimpse(ExtractLuan)
length(unique(ExtractLuan$TID_OMRS))
```