### 2020-01-22 DataWrangling/Missing Received.Rmd

---
title: "Missing Received"
output: html_document
---

```{r setup, include=FALSE}
BGD8168 <- read.delim("BGD8168List.txt", header = FALSE, sep = " ")
BGD8168 <- as.data.frame(BGD8168[, 13])
colnames(BGD8168)[1] <- "TID"
QA.BGD <- as.data.frame(QA.BGD[, 1])
colnames(QA.BGD)[1] <- "TID"
QAdf2 <- rbind(BGD8168, QA.BGD)

# in delft but not in QAdf2
stillMissing <-  subset(CAD6_delft, !(CAD6_delft$TID_Delft %in% QAdf2$TID))
# in delft but not in QA.BGD
stillMissing2 <-  subset(CAD6_delft, !(CAD6_delft$TID_Delft %in% QA.BGD$TID))
# in missing but not in BGD8168 
stillMissing3 <- subset(Missing, !(Missing$TID_OMRS %in% BGD8168$TID))
length(Missing$TID_OMRS)-length(BGD8168$TID)
```