### 2020-01-22 ROCs.Rmd

```{r Global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=10, fig.height=6, echo=FALSE, warning=FALSE, message=FALSE)

Sys.setenv(LANG = "en")
library(ggplot2)
library(scales)
library(epiR)
library(ggthemes)
library(RColorBrewer)
library(plyr)
library(reshape2)
library(pROC)
library(tidyverse)
library(dplyr)
library(lubridate)
library(readxl)
library(plotROC)
library(flextable)

# as.numeric.factor <- function(x) {as.numeric(levels(x))[x]}
```

```{r}
Adult_Valid <- read_csv('Adult_Valid.csv')
```

### Histogram 
```{r Histogram, fig.width=8, fig.height=5,}
# make it long
Adult_Valid$qXRv2_100 <- Adult_Valid$qXRv2*100

Adult_Valid_long <- gather(Adult_Valid, DeepLearningSystem, AbnormalityScore, CAD4TB6, qXRv2_100)

# p <- ggplot(Adult_Valid,aes(CAD4TB6))
# p <- ggplot(Adult_Valid,aes(qXRv2))
# p <- ggplot(Adult_Valid,aes(DrCADxScore))

Adult_Valid_long$Xpert2Outcome_num <- as.factor(Adult_Valid_long$Xpert2Outcome_num)

p <- ggplot(Adult_Valid_long,aes(x=AbnormalityScore, fill=Xpert2Outcome_num))+ geom_histogram(position = 'identity', alpha=0.5, binwidth = 10, aes(y = ..density.., fill = Xpert2Outcome_num)) + xlab("Abnormality Scores of the Deep Learning Systems") + scale_fill_discrete(name ="", labels = c("Xpert Negative", "Xpert Positive"))
p + facet_wrap(~DeepLearningSystem) + theme_minimal() + theme(legend.position = "bottom")+ scale_color_brewer(palette="Accent")

# p + facet_grid(Country ~ DeepLearningSystem, switch = "both") + theme_minimal() + theme(legend.position = "bottom")

# ggplot(Adult_Valid_long, aes(x=DL.Score, fill=DL.System)) + geom_histogram(position = 'identity', alpha = 0.5, bindwidth = 10, aes(y=..density.., fill=DL.System))

# density plot
# ggplot(Adult_Valid_long,aes(x=DL.Score, fill=Xpert2Outcome_num))+ geom_density(position = 'identity', alpha=0.5, binwidth = 10)

# pooled 
# p + geom_histogram(position = 'identity', alpha=0.5, aes(y = ..density.., fill = Xpert2Outcome_num), bins = 20) + stat_density(position = 'identity',aes(colour = Xpert2Outcome_num)) # Pooled

# library(moments)
# skewness(Adult_Valid[Adult_Valid$Xpert2Outcome_num ==1, 28]) #28 CAD4TB6, 29 qXRv2, 33 Lunit
# skewness(Adult_Valid[Adult_Valid$Xpert2Outcome_num ==1, 29]) #29 qXRv2, 33 Lunit
# skewness(Adult_Valid[Adult_Valid$Xpert2Outcome_num ==1, 33]) #33 Lunit
# 
# skewness(Adult_Valid[Adult_Valid$Xpert2Outcome_num ==0, 28]) #28 CAD4TB6, 29 qXRv2, 33 Lunit
# skewness(Adult_Valid[Adult_Valid$Xpert2Outcome_num ==0, 29]) #29 qXRv2, 33 Lunit
# skewness(Adult_Valid[Adult_Valid$Xpert2Outcome_num ==0, 33]) #33 Lunit

```

###  Multi-ROC (Xpert Reference) 
```{r Multi-ROC_All_Xpert}
# make it long
Adult_Valid$qXRv2_100 <- Adult_Valid$qXRv2*100

Adult_Valid_long <- gather(Adult_Valid, DeepLearningSystem, AbnormalityScore, CAD4TB6, qXRv2_100)

############# Pooled ###############
roc_CAD6 <- ci.auc(Xpert2Outcome_num ~ CAD4TB6, Adult_Valid)
roc_qure <- ci.auc(Xpert2Outcome_num ~ qXRv2_100, Adult_Valid)
# roc_Lunit <- ci.auc(Xpert2Outcome_num ~ Lunit, Adult_Valid)
# Adult_Valid_long$DeepLearningSystem[Adult_Valid_long$DeepLearningSystem %in% "Lunit_100"] <- "Lunit"

Adult_Valid_long$DeepLearningSystem[Adult_Valid_long$DeepLearningSystem %in% "CAD4TB6"] <- "CAD4TB"
Adult_Valid_long$DeepLearningSystem[Adult_Valid_long$DeepLearningSystem %in% "qXRv2_100"] <- "qXR"

Adult_Valid_long$DeepLearningSystem <- as.factor(Adult_Valid_long$DeepLearningSystem)

# Multi-ROC
ggROC <- ggplot(Adult_Valid_long, aes(d = Xpert2Outcome_num, m = AbnormalityScore, color = DeepLearningSystem)) + geom_roc(cutoffs.at = c(10,20,30,40,50,60,70,80,90)) + style_roc(xlab = "1 - Specificity", ylab = "Sensitivity")

ggROC <- ggROC + annotate("text", x = .75, y = .15, 
            label = paste(" CAD4TB: ", round(roc_CAD6[2],2), " (95% CI:", round(roc_CAD6[1],2), "-", round(roc_CAD6[3],2), ")", "\n",
      # "Lunit: ", round(roc_Lunit[2],2), " (95% CI:", round(roc_Lunit[1],2), "-", round(roc_Lunit[3],2), ")", "\n",
      "qXR: ", round(roc_qure[2],2), " (95% CI:", round(roc_qure[1],2), "-", round(roc_qure[3],2), ")", "\n", sep = ""), size = 4.5)+
  ggtitle("The ROC curves of CAD4TB (v6) and qXR (v2) using Xpert results as the reference (n=18682)")

ggROC
ggROC1 <- ggROC + coord_cartesian(xlim = c(0.1, 0.4), ylim = c(0.90, 1) ) +
  ggtitle("Figure 2b: Zoomed-in view of the square marked part of Fig. 2a")+theme_light()
ggROC1

# plot_interactive_roc(ggROC, width =12, height = 9.5)
# plot_interactive_roc(ggROC1, width =12, height = 4)
```

### Radiologist Accuracy (highly)
```{r Accuracy of Radiologist1}
TP <- sum(Adult_Valid$raiolgogy_highly %in% "1" & Adult_Valid$Xpert2Outcome_num==1)
FP <- sum(Adult_Valid$raiolgogy_highly %in% 1 & Adult_Valid$Xpert2Outcome_num==0)
TN <- sum(Adult_Valid$raiolgogy_highly %in% 0 & Adult_Valid$Xpert2Outcome_num==0)
FN <- sum(Adult_Valid$raiolgogy_highly %in% 0 & Adult_Valid$Xpert2Outcome_num==1)

dat <- as.table(matrix(c(TP, FP, FN, TN), nrow=2, byrow=TRUE))
colnames(dat) <- c("Xpert+","Xpert-")
rownames(dat) <- c("CXR+","CXR-")
rval <- epi.tests(dat, conf.level = 0.95)
rval
# print(rval$elements[c(47, 43,39,35)] ) 
rm(TP, FP, TN, FN)
```

### Radiologist Accuracy (possible)
```{r Accuracy of Radiologist2}
TP <- sum(Adult_Valid$raiolgogy_possible %in% "1" & Adult_Valid$Xpert2Outcome_num==1)
FP <- sum(Adult_Valid$raiolgogy_possible %in% 1 & Adult_Valid$Xpert2Outcome_num==0)
TN <- sum(Adult_Valid$raiolgogy_possible %in% 0 & Adult_Valid$Xpert2Outcome_num==0)
FN <- sum(Adult_Valid$raiolgogy_possible %in% 0 & Adult_Valid$Xpert2Outcome_num==1)

dat <- as.table(matrix(c(TP, FP, FN, TN), nrow=2, byrow=TRUE))
colnames(dat) <- c("Xpert+","Xpert-")
rownames(dat) <- c("CXR+","CXR-")
rval <- epi.tests(dat, conf.level = 0.95)
rval
# print(rval$elements[c(47, 43,39,35)] ) 
rm(TP, FP, TN, FN)
```

### Radiologist Accuracy (any abnormality)
```{r Accuracy of Radiologist3}
TP <- sum(Adult_Valid$any_abnormality %in% "1" & Adult_Valid$Xpert2Outcome_num==1)
FP <- sum(Adult_Valid$any_abnormality %in% 1 & Adult_Valid$Xpert2Outcome_num==0)
TN <- sum(Adult_Valid$any_abnormality %in% 0 & Adult_Valid$Xpert2Outcome_num==0)
FN <- sum(Adult_Valid$any_abnormality %in% 0 & Adult_Valid$Xpert2Outcome_num==1)

dat <- as.table(matrix(c(TP, FP, FN, TN), nrow=2, byrow=TRUE))
colnames(dat) <- c("Xpert+","Xpert-")
rownames(dat) <- c("CXR+","CXR-")
rval <- epi.tests(dat, conf.level = 0.95)
rval
# print(rval$elements[c(47, 43,39,35)] ) 
rm(TP, FP, TN, FN)
```

## Cutoff Table (Full)
```{r cutoff table}
###### Qure.AI Score by 0.01 #########
qure_ai_score <- seq(0.01,1, by = 0.01)

mylist <- NULL
mylist <- as.list(mylist)

for (i in 1:100){
myfunction <- function(CountryX, car.cutoff){
a <- sum(CountryX$qXRv2 >= car.cutoff & CountryX$Xpert2Outcome_num =="1")
b <- sum(CountryX$qXRv2 >= car.cutoff & CountryX$Xpert2Outcome_num =="0")
c <- sum(CountryX$qXRv2 < car.cutoff & CountryX$Xpert2Outcome_num =="1")
d <- sum(CountryX$qXRv2 < car.cutoff & CountryX$Xpert2Outcome_num =="0")
dat <- as.table(matrix(c(a,b,c,d), nrow = 2, byrow = TRUE))
colnames(dat) <- c("Dis+","Dis-")
rownames(dat) <- c("Test+","Test-")
rval <- epi.tests(dat, conf.level = 0.95)
sensitivity <- as.vector(round((rval$elements$sensitivity),2)) 
specificity <- as.vector(round((rval$elements$specificity),2))
pv.positive <- as.vector(round((rval$elements$pv.positive),2))
pv.negative <- as.vector(round((rval$elements$pv.negative),2))
correct.rate <- round((a+d)/length(CountryX$Xpert2Outcome_num), 2)

perc_above_CAR_Score<- sum(CountryX$qXRv2 >= car.cutoff)/length(CountryX$TID_OMRS)
Score <- qure_ai_score[i]

accuracy <- cbind(sensitivity, specificity, pv.positive, pv.negative, perc_above_CAR_Score, Score, correct.rate)

return(accuracy)
}

accuracy <- myfunction(Adult_Valid, qure_ai_score[i])
mylist[[i]] <- list(accuracy)

}

QureDF_both <- data.frame(matrix(unlist(mylist), nrow=100, byrow=T))
QureDF_both$Country <- paste("Pooled")

QureDF <- QureDF_both
QureDF$Sen_95CI <- paste(QureDF[, 2], "-", QureDF[, 3], sep = "")
QureDF$Spe_95CI <- paste(QureDF[, 5], "-", QureDF[, 6], sep = "")
QureDF$PPV_95CI <- paste(QureDF[, 8], "-", QureDF[, 9], sep = "")
QureDF$NPV_95CI <- paste(QureDF[, 11], "-", QureDF[, 12], sep = "")
QureDF$CAD_System <- paste("qXR")
# write.csv(QureDF, "qure.ai by 0.01.csv")
# rm(qure_ai_score)

#### CAD6 Score by 1 ######
CAD6 <- seq(1,99, by = 1)

mylist <- NULL
mylist <- as.list(mylist)

for (i in 1:99){
myfunction <- function(CountryX, car.cutoff){
a <- sum(CountryX$CAD4TB6 >= car.cutoff & CountryX$Xpert2Outcome_num =="1")
b <- sum(CountryX$CAD4TB6 >= car.cutoff & CountryX$Xpert2Outcome_num =="0")
c <- sum(CountryX$CAD4TB6 < car.cutoff & CountryX$Xpert2Outcome_num =="1")
d <- sum(CountryX$CAD4TB6 < car.cutoff & CountryX$Xpert2Outcome_num =="0")
dat <- as.table(matrix(c(a,b,c,d), nrow = 2, byrow = TRUE))
colnames(dat) <- c("Dis+","Dis-")
rownames(dat) <- c("Test+","Test-")
rval <- epi.tests(dat, conf.level = 0.95)
sensitivity <- as.vector(round((rval$elements$sensitivity),2))
specificity <- as.vector(round((rval$elements$specificity),2))
pv.positive <- as.vector(round((rval$elements$pv.positive),2))
pv.negative <- as.vector(round((rval$elements$pv.negative),2))
correct.rate <- round((a+d)/length(CountryX$Xpert2Outcome_num), 2)

perc_above_CAR_Score<- sum(CountryX$CAD4TB6 >= car.cutoff)/length(CountryX$TID_OMRS)
Score <- CAD6[i]

accuracy <- cbind(sensitivity, specificity, pv.positive, pv.negative, perc_above_CAR_Score, Score, correct.rate)

return(accuracy)
}

accuracy_both <- myfunction(Adult_Valid, CAD6[i])
mylist[[i]] <- list(accuracy_both)
}
CAD6DF_both <- data.frame(matrix(unlist(mylist), nrow=99, byrow=T))
CAD6DF_both$Country <- paste("Pooled")

CAD6DF <- CAD6DF_both
CAD6DF$Sen_95CI <- paste(CAD6DF[, 2], "-", CAD6DF[, 3], sep = "")
CAD6DF$Spe_95CI <- paste(CAD6DF[, 5], "-", CAD6DF[, 6], sep = "")
CAD6DF$PPV_95CI <- paste(CAD6DF[, 8], "-", CAD6DF[, 9], sep = "")
CAD6DF$NPV_95CI <- paste(CAD6DF[, 11], "-", CAD6DF[, 12], sep = "")
CAD6DF$CAD_System <- paste("CAD4TB")

# #### Lunit Score by 0.01  ######
# Lunit <- seq(0,1, by = 0.01)
# 
# mylist <- NULL
# mylist <- as.list(mylist)
# 
# for (i in 1:101){
# myfunction <- function(CountryX, car.cutoff){
# a <- sum(CountryX$Lunit >= car.cutoff & CountryX$Xpert2Outcome_num =="1")
# b <- sum(CountryX$Lunit >= car.cutoff & CountryX$Xpert2Outcome_num =="0")
# c <- sum(CountryX$Lunit < car.cutoff & CountryX$Xpert2Outcome_num =="1")
# d <- sum(CountryX$Lunit < car.cutoff & CountryX$Xpert2Outcome_num =="0")
# dat <- as.table(matrix(c(a,b,c,d), nrow = 2, byrow = TRUE))
# colnames(dat) <- c("Dis+","Dis-")
# rownames(dat) <- c("Test+","Test-")
# rval <- epi.tests(dat, conf.level = 0.95)
# sensitivity <- as.vector(round((rval$elements$sensitivity),2))
# specificity <- as.vector(round((rval$elements$specificity),2))
# pv.positive <- as.vector(round((rval$elements$pv.positive),2))
# pv.negative <- as.vector(round((rval$elements$pv.negative),2))
# correct.rate <- round((a+d)/length(CountryX$Xpert2Outcome_num), 2)
# 
# 
# perc_above_CAR_Score<- sum(CountryX$Lunit >= car.cutoff)/length(CountryX$TID_OMRS)
# Score <- Lunit[i]
# 
# accuracy <- cbind(sensitivity, specificity, pv.positive, pv.negative, perc_above_CAR_Score, Score, correct.rate)
# 
# return(accuracy)
# }
# 
# accuracy_both <- myfunction(Adult_Valid, Lunit[i])
# mylist[[i]] <- list(accuracy_both)
# }
# 
# LunitDF_both <- data.frame(matrix(unlist(mylist), nrow=101, byrow=T))
# LunitDF_both$Country <- paste("Pooled")
# 
# LunitDF <- LunitDF_both
# LunitDF$Sen_95CI <- paste(LunitDF[, 2], "-", LunitDF[, 3], sep = "")
# LunitDF$Spe_95CI <- paste(LunitDF[, 5], "-", LunitDF[, 6], sep = "")
# LunitDF$PPV_95CI <- paste(LunitDF[, 8], "-", LunitDF[, 9], sep = "")
# LunitDF$NPV_95CI <- paste(LunitDF[, 11], "-", LunitDF[, 12], sep = "")
# LunitDF$CAD_System <- paste("Lunit")

######### Merge DFs ######
CAD_Xpert <- rbind(CAD6DF, QureDF)
colnames(CAD_Xpert)[1] <- "Sens"
colnames(CAD_Xpert)[2] <- "Sens_L"
colnames(CAD_Xpert)[3] <- "Sens_H"
colnames(CAD_Xpert)[4] <- "Spec"
colnames(CAD_Xpert)[5] <- "Spec_L"
colnames(CAD_Xpert)[6] <- "Spec_H"
colnames(CAD_Xpert)[7] <- "ppv"
colnames(CAD_Xpert)[8] <- "PPV_L"
colnames(CAD_Xpert)[9] <- "PPV_H"
colnames(CAD_Xpert)[10] <- "npv"
colnames(CAD_Xpert)[11] <- "NPV_L"
colnames(CAD_Xpert)[12] <- "NPV_H"
colnames(CAD_Xpert)[13] <- "% Above"
colnames(CAD_Xpert)[14] <- "Score"
colnames(CAD_Xpert)[15] <- "Correct.rate"
CAD_Xpert$Sensitivity <- paste(CAD_Xpert$Sens, "(", CAD_Xpert$Sen_95CI, ")", sep = "")
CAD_Xpert$Specificity <- paste(CAD_Xpert$Spec, "(", CAD_Xpert$Spe_95CI, ")", sep = "")
CAD_Xpert$PPV <- paste(CAD_Xpert$ppv, "(", CAD_Xpert$PPV_95CI, ")", sep = "")
CAD_Xpert$NPV <- paste(CAD_Xpert$npv, "(", CAD_Xpert$NPV_95CI, ")", sep = "")
CAD_Xpert$`% Above` <- round(CAD_Xpert$`% Above`, 2)

CAD_Xpert <- CAD_Xpert[, c(16, 21, 14, 13, 15, 22:25)]

# write.csv(CAD_Xpert, "Supp_table.csv")
```