### 2020-01-22 Rename.Rmd

# Clean the wrong file name with DICOM tag

```{r Import DICOM Ex}
###### Match by filename
DICOM_extract <- read_excel("DICOM extract.xlsx")

LabelbyFileName <- merge(DICOM_extract, MDF, by.x = 'File name', by.y = 'TID_OMRS', all.x = T)

View(LabelbyFileName[, c(1,2,3,4,5,6,7,12,13)])
LabelbyFileName$Age <- as.numeric(LabelbyFileName$Age)
LabelbyFileName$AgeDICOM <- as.numeric(LabelbyFileName$AgeDICOM)

LabelbyFileName$agediff <- LabelbyFileName$Age - LabelbyFileName$AgeDICOM
summary(LabelbyFileName$agediff)
hist(LabelbyFileName$agediff, breaks = 100)

sum(LabelbyFileName$agediff %in% "0")

sum(LabelbyFileName$Gender %in%  LabelbyFileName$PatientSex & LabelbyFileName$Match %in% "FALSE")

###### Match by DICOM Tag
LabelbyDICOMName <- merge(DICOM_extract, MDF, by.x = 'PatientID', by.y = 'TID_OMRS', all.x = T)

View(LabelbyDICOMName[, c(1,2,3,4,5,6,7,12,13)])
LabelbyDICOMName$Age <- as.numeric(LabelbyDICOMName$Age)
LabelbyDICOMName$AgeDICOM <- as.numeric(LabelbyDICOMName$AgeDICOM)

LabelbyDICOMName$agediff <- LabelbyDICOMName$Age - LabelbyDICOMName$AgeDICOM
summary(LabelbyDICOMName$agediff)
hist(LabelbyDICOMName$agediff, breaks = 100)

sum(LabelbyDICOMName$agediff %in% "0")

sum(LabelbyDICOMName$Gender %in% LabelbyDICOMName$PatientSex & LabelbyDICOMName$Match %in% "FALSE")

#### 4 Luan #####
###### Match by filename

DICOMTag_Luan <- read_excel("DICOMTag Luan.xlsx")
LabelbyFileName <- merge(DICOMTag_Luan, MDF_Delft, by.x = 'FileName', by.y = 'TID_OMRS', all.x = T)

View(LabelbyFileName[, c(1,2,3,4,5,6,7,8, 14, 15)])
summary(LabelbyFileName$Match)

LabelbyFileName$Age <- as.numeric(LabelbyFileName$Age)
LabelbyFileName$AgeDICOM <- as.numeric(LabelbyFileName$AgeDICOM)

LabelbyFileName$agediff <- LabelbyFileName$Age - LabelbyFileName$AgeDICOM
summary(LabelbyFileName$agediff)
hist(LabelbyFileName$agediff, breaks = 100)

sum(LabelbyFileName$agediff %in% "0")

sum(LabelbyFileName$Gender %in%  LabelbyFileName$PatientSex & LabelbyFileName$Match %in% "FALSE")

###### Match by DICOM Tag
LabelbyDICOMName <- merge(DICOM_extract, MDF, by.x = 'PatientID', by.y = 'TID_OMRS', all.x = T)

View(LabelbyDICOMName[, c(1,2,3,4,5,6,7,12,13)])
LabelbyDICOMName$Age <- as.numeric(LabelbyDICOMName$Age)
LabelbyDICOMName$AgeDICOM <- as.numeric(LabelbyDICOMName$AgeDICOM)

LabelbyDICOMName$agediff <- LabelbyDICOMName$Age - LabelbyDICOMName$AgeDICOM
summary(LabelbyDICOMName$agediff)
hist(LabelbyDICOMName$agediff, breaks = 100)

sum(LabelbyDICOMName$agediff %in% "0")

sum(LabelbyDICOMName$Gender %in% LabelbyDICOMName$PatientSex & LabelbyDICOMName$Match %in% "FALSE")

```

```{r Cleaning Validated}
validated <- read_csv("validated.csv")
n_occur <- data.frame(table(validated$ValidedTID))
table(n_occur$Freq)
validated <- as.data.frame(n_occur$Var1[n_occur$Freq<2])
colnames(validated)[1] <- "TID"
n_occur <- data.frame(table(validated$TID))
table(n_occur$Freq)

validated_merge <- merge(validated, Master_df, by.x="TID", by.y = "TID_OMRS")
n_occur <- data.frame(table(validated_merge$TID))
table(n_occur$Freq)
validated_merge <- as.data.frame(n_occur$Var1[n_occur$Freq<2])
colnames(validated_merge)[1] <- "TID"
write.csv(validated_merge, "corrected_merge.csv")

missing <- subset(Master_df, !(Master_df$TID_OMRS %in% validated_merge$TID))
# write.csv(missing, "missing.csv")
transferred <- read_excel("transferred.xlsx")
missing <- subset(validated_merge, !(validated_merge$TID %in% transferred$Transferred))

write.csv(missing, "missing.csv")

rm(n_occur)
```