### 2020-07-28 Chapter/Tele.Rmd
CAD6.3

```{r}
library(readxl)
Teleradiology <- read_excel("C:/Users/zhizh/OneDrive - Stop TB Partnership/UNOPS/10 Paper Writing/CAR software/02 Bangladesh/MachineBGD/DataWrangling/TeleradiologyReportsUpdated.xlsx")
# names(Teleradiology)
colnames(Teleradiology)[1] <- "TID_OMRS"

Teleradiology$`Abnormality/TB`[Teleradiology$`Abnormality/TB` == "Abnormality consistent with TB" ] <- "TB_Abnormality"
Teleradiology$`Abnormality/TB`[Teleradiology$`Abnormality/TB` == "Abnormality NOT consistent with TB" ] <- "All_Abnormality"

Teleradiology[!Teleradiology$TID_OMRS %in% Tele$TID_OMRS, 1]

Tele <- merge(Teleradiology, MDF[, c(1, 3, 13, 46:52)], by="TID_OMRS")

# write.csv(Tele, "Tele.csv")
t1 <- table(Tele$Comment, Tele$`Abnormality/TB`)
# ftable(addmargins(t1))
# t1[4, ]  <- t1[4, ] +t1[6, ]
# t1[2, ]  <- t1[2, ] +t1[4, ]
# 
# t1[7, ]  <- t1[7, ] +t1[8, ]
# t1[5, ]  <- t1[5, ] +t1[7, ]
# t1[3, ]  <- t1[3, ] +t1[5, ]

t1 <- t1[c(2,4,6,8,10, 3,5,7,9,11, 12), ]
addmargins(t1)
# teleTable <- as.data.frame()
teleTable <- matrix(addmargins(t1), nrow = 12, ncol=4, byrow = FALSE)

row.names(teleTable) <- c("Missed by 1 AI AND by human" ,  "Missed by 2 AI AND human ", "Missed by 3 AI AND human ", "Missed by 4 AI AND human ",  "Missed by 5 AI AND human ", "Missed by 1 AI but NOT human",  "Missed by 2 AI but NOT human ", "Missed by 3 AI but NOT human ", "Missed by 4 AI but NOT human ", "Missed by 5 AI but NOT human", "Missed by human ONLY ", "Total")
colnames(teleTable) <- c("All_Abnormality", "Normal", "TB_Abnormality", "Total")

teleTable <- as.data.frame(teleTable)

MarginTable <- percent(matrix(prop.table(t1, margin = 1), nrow = 11, ncol=3, byrow = FALSE))
MarginTable <- matrix(MarginTable, nrow = 11, ncol=3, byrow = FALSE)
row.names(MarginTable) <- c("Missed by 1 AI AND by human" ,  "Missed by 2 AI AND human ", "Missed by 3 AI AND human ", "Missed by 4 AI AND human ",  "Missed by 5 AI AND human ", "Missed by 1 AI but NOT human",  "Missed by 2 AI but NOT human ", "Missed by 3 AI but NOT human ", "Missed by 4 AI but NOT human ", "Missed by 5 AI but NOT human", "Missed by human ONLY ")

colnames(MarginTable) <- c("All_Abnormality", "Normal", "TB_Abnormality")
MarginTable <- as.data.frame(MarginTable)

MarginTable[nrow(MarginTable)+1, ] = c("", "", "")
row.names(MarginTable)[12] <- "Total"

TeleTable <- cbind(teleTable, MarginTable)
rm(teleTable, MarginTable)

TeleTable <- TeleTable[, c(3,7,1,5,2,6,4)]
TeleTable
# write.csv(TeleTable, "TeleTable.csv")
```

## Rradiologist Accuracy
```{r Rradiologist Accuracy}
MDF$USHuman <- Tele$`Abnormality/TB`[match(MDF$TID_OMRS, Tele$TID_OMRS)]
USOnly <- MDF[is.na(MDF$USHuman)==FALSE, ]

## BGD radiologists
  ###  High+possible ----------------------------
  TP <- sum(USOnly$rad.TB %in% "1" & USOnly$Xpert2Outcome_num==1)
  FP <- sum(USOnly$rad.TB %in% 1 & USOnly$Xpert2Outcome_num==0)
  TN <- sum(USOnly$rad.TB %in% 0 & USOnly$Xpert2Outcome_num==0)
  FN <- sum(USOnly$rad.TB %in% 0 & USOnly$Xpert2Outcome_num==1)

  dat <- as.table(matrix(c(TP, FP, FN, TN), nrow=2, byrow=TRUE))
  colnames(dat) <- c("Xpert+","Xpert-")
  rownames(dat) <- c("CXR+","CXR-")
  rval <- epi.tests(dat, conf.level = 0.95)
  Probably <- rval$elements[c(59, 55)]
  rm(TP, FP, TN, FN)

  ### Any Abnormalities ------------------------------
  TP <- sum(USOnly$rad.abn %in% "1" & USOnly$Xpert2Outcome_num==1)
  FP <- sum(USOnly$rad.abn %in% 1 & USOnly$Xpert2Outcome_num==0)
  TN <- sum(USOnly$rad.abn %in% 0 & USOnly$Xpert2Outcome_num==0)
  FN <- sum(USOnly$rad.abn %in% 0 & USOnly$Xpert2Outcome_num==1)

  dat <- as.table(matrix(c(TP, FP, FN, TN), nrow=2, byrow=TRUE))
  colnames(dat) <- c("Xpert+","Xpert-")
  rownames(dat) <- c("CXR+","CXR-")
  rval <- epi.tests(dat, conf.level = 0.95)
  Any <- rval$elements[c(59, 55)]
  rm(TP, FP, TN, FN)

  BGDReader <- data.frame(matrix(unlist(c(Probably, Any)), nrow = 2, byrow = T))
  colnames(BGDReader) <- c("Sens", "Sens_L", "Sens_H", "Spec", "Spec_L", "Spec_H")
  BGDReader$X <- 1-BGDReader$Spec
  BGDReader$BGDReaderCategory <- c("Probably TB", "Any Abnormality")

## US radiologists
###  TB Abnormality ----------------------------
  TP <- sum(USOnly$USHuman %in% "TB_Abnormality" & USOnly$Xpert2Outcome_num==1)
  FP <- sum(USOnly$USHuman %in% "TB_Abnormality" & USOnly$Xpert2Outcome_num==0)
  TN <- sum(USOnly$USHuman != "TB_Abnormality" & USOnly$Xpert2Outcome_num==0)
  FN <- sum(USOnly$USHuman != "TB_Abnormality" & USOnly$Xpert2Outcome_num==1)

  dat <- as.table(matrix(c(TP, FP, FN, TN), nrow=2, byrow=TRUE))
  colnames(dat) <- c("Xpert+","Xpert-")
  rownames(dat) <- c("CXR+","CXR-")
  rval <- epi.tests(dat, conf.level = 0.95)
  TB <- rval$elements[c(59, 55)]
  rm(TP, FP, TN, FN)

###  All Abnormality ----------------------------
  TP <- sum(USOnly$USHuman != "Normal" & USOnly$Xpert2Outcome_num==1)
  FP <- sum(USOnly$USHuman != "Normal" & USOnly$Xpert2Outcome_num==0)
  TN <- sum(USOnly$USHuman == "Normal" & USOnly$Xpert2Outcome_num==0)
  FN <- sum(USOnly$USHuman == "Normal" & USOnly$Xpert2Outcome_num==1)

  dat <- as.table(matrix(c(TP, FP, FN, TN), nrow=2, byrow=TRUE))
  colnames(dat) <- c("Xpert+","Xpert-")
  rownames(dat) <- c("CXR+","CXR-")
  rval <- epi.tests(dat, conf.level = 0.95)
  All <- rval$elements[c(59, 55)]

  USReader <- data.frame(matrix(unlist(c(TB, All)), nrow = 2, byrow = T))
  colnames(USReader) <- c("Sens", "Sens_L", "Sens_H", "Spec", "Spec_L", "Spec_H")
  USReader$X <- 1-USReader$Spec
  rm(TP, FP, TN, FN)

USReader
BGDReader
```

## ROC
```{r ROC}
library(ggplot2)
library(scales)
library(epiR)
library(ggthemes)
library(RColorBrewer)
library(plyr)
library(reshape2)
library(pROC)
library(tidyverse)
library(dplyr)
library(lubridate)
library(readxl)
library(plotROC)
library(flextable)
library(DataExplorer)

roc_CAD6 <- ci.auc(Xpert2Outcome_num ~ CAD4TB6, USOnly)
roc_qure <- ci.auc(Xpert2Outcome_num ~ qXRv3_100, USOnly)
roc_Lunit <- ci.auc(Xpert2Outcome_num ~ LunitScore_100, USOnly)
roc_JF1 <- ci.auc(Xpert2Outcome_num ~ JF1_100, USOnly)
roc_JF2 <- ci.auc(Xpert2Outcome_num ~ JF2_100, USOnly)
roc_IF1 <- ci.auc(Xpert2Outcome_num ~ IF1_100, USOnly)
roc_IF2 <- ci.auc(Xpert2Outcome_num ~ IF2_100, USOnly)
roc_IF3 <- ci.auc(Xpert2Outcome_num ~ IF3_100, USOnly)

USOnly_long <- gather(USOnly, DeepLearningSystem, AbnormalityScore, CAD4TB6, qXRv3_100, LunitScore_100, JF1_100, IF2_100)

ggROC <- ggplot(USOnly_long, aes(d = Xpert2Outcome_num, m = AbnormalityScore, color = DeepLearningSystem)) + geom_roc(cutoffs.at = c(50,80)) + style_roc(xlab = "1 - Specificity", ylab = "Sensitivity")

ggROC <- ggROC + annotate("text", x = .75, y = .15, label = paste("Lunit INSIGHT CXR: ", round(roc_Lunit[2],2), " (95% CI:", round(roc_Lunit[1],2), "-", round(roc_Lunit[3],2), ")", "\n", "qXR: ", round(roc_qure[2],2), " (95% CI:", round(roc_qure[1],2), "-", round(roc_qure[3],2), ")", "\n", "InferReadDR: ", round(roc_IF2[2],2), " (95% CI:", round(roc_IF2[1],2), "-", round(roc_IF2[3],2), ")", "\n", "JF CXR-1: ", round(roc_JF1[2],2), " (95% CI:", round(roc_JF1[1],2), "-", round(roc_JF1[3],2), ")", "\n",  "CAD4TB: ", round(roc_CAD6[2],2), " (95% CI:", round(roc_CAD6[1],2), "-", round(roc_CAD6[3],2), ")", sep = ""), size = 3)
# 
# ggROC <- ggROC + geom_hline(yintercept = 0.8096192, linetype="dotted", color = "blue", size=0.5) + geom_vline(xintercept = 0.2105263, linetype="dotted", color = "blue", size=0.5)
# ggROC <- ggROC + geom_hline(yintercept = 0.7134269, linetype="dotted", color = "red", size=0.5) + geom_vline(xintercept = 0.1578947, linetype="dotted", color = "red", size=0.5)

ggROC <- ggROC + geom_hline(yintercept =  0.5030060, linetype="dotted", color = "blue", size=0.5) + geom_vline(xintercept = 0.1578947, linetype="dotted", color = "blue", size=0.5)
ggROC <- ggROC + geom_hline(yintercept = 0.8096192, linetype="dotted", color = "red", size=0.5) + geom_vline(xintercept = 0.2105263, linetype="dotted", color = "red", size=0.5)

ggROC
```