### 2021-01-15 SubGroup.Rmd
lancet

```{r Global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=15, fig.height=12, echo=FALSE, warning=FALSE, message=FALSE)
```

```{r eval=F}
source("DataWrangling/GlobalOption.R")
source("radiologist.R")
Radiologist <- Radiologist[Radiologist$Referral %in% "MDF", ]

New <- MDF[MDF$TB.Medication.History %in% "No",]
Old <- MDF[MDF$TB.Medication.History %in% "Yes",]

Female <- MDF[MDF$Gender %in% "F", ]
Male <- MDF[MDF$Gender %in% "M", ]

Young <- MDF[MDF$AgeGroup %in% "[15,25)",]
Middle <- MDF[MDF$AgeGroup %in% "[25,60)",]
Senior <- MDF[MDF$AgeGroup %in% "[60,108]",]

PrivateReferral <-subset(MDF, MDF$UseCase %in% "PrivateReferral")
PublicReferral <-subset(MDF, MDF$UseCase %in% "PublicReferral")
DOTS <-subset(MDF, MDF$UseCase %in% "PublicDOTSRetesting")
WalkIn <-subset(MDF, MDF$UseCase %in% "WalkIn")
Community <-subset(MDF, MDF$UseCase %in% "Community screening")
Contacts <-subset(MDF, MDF$UseCase %in% "Contacts")
# Missing <- subset(MDF, is.na(MDF$UseCase)==T)

New_long <- MDF_long[MDF_long$TB.Medication.History %in% "No",]
Old_long <- MDF_long[MDF_long$TB.Medication.History %in% "Yes",]
Young_long <- MDF_long[MDF_long$AgeGroup %in% "[15,25)",]
Middle_long <- MDF_long[MDF_long$AgeGroup %in% "[25,60)",]
Senior_long <- MDF_long[MDF_long$AgeGroup %in% "[60,108]",]

PrivateReferral_long <-subset(MDF_long, MDF_long$UseCase %in% "PrivateReferral")
PublicReferral_long <-subset(MDF_long, MDF_long$UseCase %in% "PublicReferral")
DOTS_long <-subset(MDF_long, MDF_long$UseCase %in% "PublicDOTSRetesting")
WalkIn_long <-subset(MDF_long, MDF_long$UseCase %in% "WalkIn")
Community_long <-subset(MDF_long, MDF_long$UseCase %in% "Community screening")
Contacts_long <-subset(MDF_long, MDF_long$UseCase %in% "Contacts")
# Missing <- subset(MDF_long, is.na(MDF_long$UseCase)==T)

```

```{r 2 AUC}
## All
library(precrec)
library(ggplot2)

ROCPRC <- function(dataset){
  attr <- attributes(evalmod(scores = dataset$CAD4TB6, labels = dataset$Xpert2Outcome_num))
  aucT1 <- c("CAD4TB ", round(attr$auc[4],3))

  attr <- attributes(evalmod(scores = dataset$qXRv3, labels = dataset$Xpert2Outcome_num))
  aucT2 <- c("qXR ", round(attr$auc[4],3))

  attr <- attributes(evalmod(scores = dataset$LunitScore, labels = dataset$Xpert2Outcome_num))
  aucT3 <- c("Lunit INSIGHT CXR", round(attr$auc[4],3))

  attr <- attributes(evalmod(scores = dataset$JF1, labels = dataset$Xpert2Outcome_num))
  aucT4 <- c("JF CXR-1", round(attr$auc[4],3))

  attr <- attributes(evalmod(scores = dataset$IF2, labels = dataset$Xpert2Outcome_num))
  aucT5 <- c("InferReadDR ", round(attr$auc[4],3))

  # return(mmcurves)
  # return(autoplot(mmcurves))
  aucT <- c(aucT1, aucT2, aucT3, aucT4, aucT5)
  return(aucT)
}

l <- ROCPRC(MDF)
all <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
all$subgroup <- "all"

## New only
l <- ROCPRC(New)
New <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
New$subgroup <- "New cases"

## Old on ly
l <- ROCPRC(Old)
Old <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
Old$subgroup <- "Previously treated cases"

## Young
l <- ROCPRC(Young)
Young <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
Young$subgroup <- "Young age"

## Middle
l <- ROCPRC(Middle)
Middle <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
Middle$subgroup <- "Middle age"

## Senior
l <- ROCPRC(Senior)
Senior <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
Senior$subgroup <- "Old age"

## PrivateReferral only
l <- ROCPRC(PrivateReferral)
PrivateReferral <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
PrivateReferral$subgroup <- "PrivateReferral"

## PublicReferral only
l <- ROCPRC(PublicReferral)
PublicReferral <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
PublicReferral$subgroup <- "PublicReferral"

## DOTS only
l <- ROCPRC(DOTS)
DOTS <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
DOTS$subgroup <- "DOTS retested"

## Walk-in only
l <- ROCPRC(WalkIn)
WalkIn <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
WalkIn$subgroup <- "WalkIn"

## Community only
l <- ROCPRC(Community)
Community <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
Community$subgroup <- "Community"

## Contacts only
l <- ROCPRC(Contacts)
Contacts <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
Contacts$subgroup <- "Contacts"

## Female
l <- ROCPRC(Female)
Female <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
Female$subgroup <- "Female"

## Male
l <- ROCPRC(Male)
Male <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
Male$subgroup <- "Male"
#### stratification by age AND prior TB history

aucTable <- rbind(all, New, Old, Young, Middle, Senior, PrivateReferral, PublicReferral, DOTS, WalkIn,Community, Contacts, Female, Male)

colnames(aucTable) <- c("AI.Algorithm", "AUC", "PCAUC", "Subgroup")
write.csv(aucTable, "Results/aucTable.csv")
rm(list=(ls()))
```

```{r ROC AUC with CI}

library(pROC)
ROC <- function(dataset){
  roc_CAD6 <- ci.auc(Xpert2Outcome_num ~ CAD4TB6, dataset)
  roc_qure <- ci.auc(Xpert2Outcome_num ~ qXRv3_100, dataset)
  roc_Lunit <- ci.auc(Xpert2Outcome_num ~ LunitScore_100, dataset)
  roc_JF1 <- ci.auc(Xpert2Outcome_num ~ JF1_100, dataset)
  roc_IF2 <- ci.auc(Xpert2Outcome_num ~ IF2_100, dataset)

  aucT <- as.data.frame(matrix(c(roc_CAD6[2], roc_CAD6[1], roc_CAD6[3], roc_qure[2], roc_qure[1], roc_qure[3], roc_Lunit[2], roc_Lunit[1], roc_Lunit[3], roc_JF1[2], roc_JF1[1], roc_JF1[3], roc_IF2[2], roc_IF2[1], roc_IF2[3]),  byrow=TRUE, ncol =  3))
  names(aucT) <- c("AUC", "AUCL", "AUCH")
  aucT$AI.Algorithm <- c("CAD4TB", "qXR", "Lunit INSIGHT CXR", "JF CXR-1", "InferReadDR")

  return(aucT)
}

## New only
New <- ROC(New)
New$subgroup <- "New cases"

## Old only
Old <- ROC(Old)
Old$subgroup <- "Previously treated cases"

## Young
Young <- ROC(Young)
Young$subgroup <- "Young age"

## Middle
Middle <- ROC(Middle)
Middle$subgroup <- "Middle age"

## Senior
Senior <- ROC(Senior)
Senior$subgroup <- "Old age"

## Referral only
Referral <- ROC(Referral)
Referral$subgroup <- "Referred"

## DOTS only
DOTS <- ROC(DOTS)
DOTS$subgroup <- "DOTS retested"

## Walk-in only
WalkIn <- ROC(WalkIn)
WalkIn$subgroup <- "WalkIn"

ROCaucTable <- rbind(New, Old, Young, Middle, Senior, Referral, DOTS, WalkIn)
View(ROCaucTable)
write.csv(ROCaucTable, "Results/ROCaucTable.csv", row.names = F)
```

```{r Bar chart}
aucTable <- read.csv("Results/aucTable.csv")
aucTable <- aucTable[, -1]

aucTable$Subgroup <- factor(aucTable$Subgroup, levels = c("All","Young age", "Middle age", "Old age", "PrivateReferral", "PublicReferral", "DOTS", "WalkIn", "Community", "Contacts", "New cases", "Previously treated cases", "Female", "Male"))
### AUC
age<-ggplot(aucTable[aucTable$Subgroup %in% c("Young age", "Middle age", "Old age"), ], aes(x=AI.Algorithm, y=AUC, fill=Subgroup)) +  geom_bar(stat="identity", width=0.6, position=position_dodge())+theme_minimal()+ theme(legend.position = "top")+  ylim(0, 0.95) + theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.95, 0.05), breaks = seq(0.65, 0.95, 0.05))

referral<-ggplot(aucTable[aucTable$Subgroup %in% c("PrivateReferral", "PublicReferral", "DOTS", "WalkIn", "Community", "Contacts"), ], aes(x=AI.Algorithm, y=AUC, fill=Subgroup)) + geom_bar(stat="identity", width=0.7, position=position_dodge())+theme_minimal()+ theme(legend.position = "top")+  ylim(0, 0.95) + theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.95, 0.05), breaks = seq(0.65, 0.95, 0.05))

history<-ggplot(aucTable[aucTable$Subgroup %in% c("New cases", "Previously treated cases"), ], aes(x=AI.Algorithm, y=AUC, fill=Subgroup)) + geom_bar(stat="identity", width=0.5, position=position_dodge())+theme_minimal() + theme(legend.position = "top") +  ylim(0, 0.95) + theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.95, 0.05), breaks = seq(0.65, 0.95, 0.05))

gender <- ggplot(aucTable[aucTable$Subgroup %in% c("Female", "Male"), ], aes(x=AI.Algorithm, y=AUC, fill=Subgroup)) + geom_bar(stat="identity", width=0.5, position=position_dodge())+theme_minimal() + theme(legend.position = "top") +  ylim(0, 0.95) + theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.95, 0.05), breaks = seq(0.65, 0.95, 0.05))

### PRC
agePC<-ggplot(aucTable[aucTable$Subgroup %in% c("Young age", "Middle age", "Old age"), ], aes(x=AI.Algorithm, y=PCAUC, fill=Subgroup)) +  geom_bar(stat="identity", width=0.6, position=position_dodge())+theme_minimal()+ theme(legend.position = "top")+  ylim(0.3, 0.8) + theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.8, 0.05), breaks = seq(0.3, 0.8, 0.05))

referralPC<-ggplot(aucTable[aucTable$Subgroup %in% c("PrivateReferral", "PublicReferral", "DOTS", "WalkIn", "Community", "Contacts"), ], aes(x=AI.Algorithm, y=PCAUC, fill=Subgroup)) + geom_bar(stat="identity", width=0.7, position=position_dodge())+theme_minimal()+ theme(legend.position = "top")+  ylim(0.3, 0.8) + theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.8, 0.05), breaks = seq(0.3, 0.8, 0.05))

historyPC<-ggplot(aucTable[aucTable$Subgroup %in% c("New cases", "Previously treated cases"), ], aes(x=AI.Algorithm, y=PCAUC, fill=Subgroup)) + geom_bar(stat="identity", width=0.5, position=position_dodge())+theme_minimal() + theme(legend.position = "top") +  ylim(0.3, 0.8) + theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.8, 0.05), breaks = seq(0.3, 0.8, 0.05))

genderPC <- ggplot(aucTable[aucTable$Subgroup %in% c("Female", "Male"), ], aes(x=AI.Algorithm, y=PCAUC, fill=Subgroup)) + geom_bar(stat="identity", width=0.5, position=position_dodge())+theme_minimal() + theme(legend.position = "top") +  ylim(0, 0.95) + theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.95, 0.05), breaks = seq(0.65, 0.95, 0.05))

tiff("Results/PRC ROC AUC plot 3.tif", width = 20, height = 8, units = "in", res = 110)
require(gridExtra)
grid.arrange(age, referral, history, gender, agePC, referralPC, historyPC, genderPC, ncol=4)
dev.off()

# rm(age, referral, history, agePC, referralPC, historyPC, gender, genderPC)

tiff("Results/PRC AUC plot.tif", width = 9, height = 8, units = "in", res = 100)
grid.arrange(agePC, referralPC, historyPC, genderPC, ncol=2)
dev.off()
```

```{r}
ROCaucTable <- read.csv("Results/ROCaucTable.csv")
ROCaucTable$subgroup <- factor(ROCaucTable$subgroup, levels = c("Young age", "Middle age", "Old age", "Referred", "DOTS retested", "WalkIn", "New cases", "Previously treated cases"))
### AUC
age <-ggplot(ROCaucTable[ROCaucTable$subgroup %in% c("Young age", "Middle age", "Old age"), ], aes(x=AI.Algorithm, y=AUC, fill=subgroup)) +  geom_bar(stat="identity", width=0.6, position=position_dodge()) + geom_errorbar(aes(ymin=AUCL, ymax=AUCH), width=.2, position=position_dodge(0.7)) + theme_minimal()+ theme(legend.position = "top")+  ylim(0, 0.95) + theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.95, 0.05), breaks = seq(0.65, 0.95, 0.05)) 
age

referral<-ggplot(ROCaucTable[ROCaucTable$subgroup %in% c("Referred", "DOTS retested", "WalkIn"), ], aes(x=AI.Algorithm, y=AUC, fill=subgroup)) + geom_bar(stat="identity", width=0.7, position=position_dodge())+theme_minimal()+ theme(legend.position = "top")+  ylim(0, 0.95) + theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.95, 0.05), breaks = seq(0.65, 0.95, 0.05)) + geom_errorbar(aes(ymin=AUCL, ymax=AUCH), width=.2, position=position_dodge(0.6)) 
referral

history<-ggplot(ROCaucTable[ROCaucTable$subgroup %in% c("New cases", "Previously treated cases"), ], aes(x=AI.Algorithm, y=AUC, fill=subgroup)) + geom_bar(stat="identity", width=0.5, position=position_dodge())+theme_minimal() + theme(legend.position = "top") +  ylim(0, 0.95) + theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.95, 0.05), breaks = seq(0.65, 0.95, 0.05)) + geom_errorbar(aes(ymin=AUCL, ymax=AUCH), width=.2, position=position_dodge(0.6))

tiff("Results/ROC_AUC plot.tif", width = 13, height = 3, units = "in", res = 100)
require(gridExtra)

grid.arrange(age, referral, history, ncol=3)

dev.off()
rm(age, referral, history)
```