### 2021-06-24 SubGroup.Rmd
lancet publication

```{r Global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=15, fig.height=12, echo=FALSE, warning=FALSE, message=FALSE)
```

```{r run manually eval=F}
rm(list = ls())

source("DataWrangling/GlobalOption.R")
source("radiologist.R")
Radiologist <- Radiologist[Radiologist$Referral %in% "MDF", ]

New <- MDF[MDF$TB.Medication.History %in% "No",]
Old <- MDF[MDF$TB.Medication.History %in% "Yes",]

Female <- MDF[MDF$Gender %in% "F", ]
Male <- MDF[MDF$Gender %in% "M", ]

Young <- MDF[MDF$AgeGroup %in% "[15,25)",]
Middle <- MDF[MDF$AgeGroup %in% "[25,60)",]
Senior <- MDF[MDF$AgeGroup %in% "[60,108]",]

PrivateReferral <-subset(MDF, MDF$UseCase %in% "PrivateReferral")
PublicReferral <-subset(MDF, MDF$UseCase %in% "PublicReferral")
DOTS <-subset(MDF, MDF$UseCase %in% "PublicDOTSRetesting")
WalkIn <-subset(MDF, MDF$UseCase %in% "WalkIn")
Community <-subset(MDF, MDF$UseCase %in% "Community screening")
Contacts <-subset(MDF, MDF$UseCase %in% "Contacts")
```

```{r 2 AUCs in subgroups}
## All
library(precrec)
library(ggplot2)

ROCPRC <- function(dataset){
  attr <- attributes(evalmod(scores = dataset$CAD4TB6, labels = dataset$Xpert2Outcome_num))
  aucT1 <- c("CAD4TB ", round(attr$auc[4],4))

  attr <- attributes(evalmod(scores = dataset$qXRv3, labels = dataset$Xpert2Outcome_num))
  aucT2 <- c("qXR ", round(attr$auc[4],4))

  attr <- attributes(evalmod(scores = dataset$LunitScore, labels = dataset$Xpert2Outcome_num))
  aucT3 <- c("Lunit INSIGHT CXR", round(attr$auc[4],4))

  attr <- attributes(evalmod(scores = dataset$JF1, labels = dataset$Xpert2Outcome_num))
  aucT4 <- c("JF CXR-1", round(attr$auc[4],4))

  attr <- attributes(evalmod(scores = dataset$IF2, labels = dataset$Xpert2Outcome_num))
  aucT5 <- c("InferReadDR ", round(attr$auc[4],4))

  # return(mmcurves)
  # return(autoplot(mmcurves))
  aucT <- c(aucT1, aucT2, aucT3, aucT4, aucT5)
  return(aucT)
}

l <- ROCPRC(MDF)
all <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
all$subgroup <- "all"

## New only
l <- ROCPRC(New)
New <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
New$subgroup <- "New cases"

## Old only
l <- ROCPRC(Old)
Old <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
Old$subgroup <- "Previously treated cases"

## Young
l <- ROCPRC(Young)
Young <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
Young$subgroup <- "Young age"

## Middle
l <- ROCPRC(Middle)
Middle <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
Middle$subgroup <- "Middle age"

## Senior
l <- ROCPRC(Senior)
Senior <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
Senior$subgroup <- "Old age"

## PrivateReferral only
l <- ROCPRC(PrivateReferral)
PrivateReferral <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
PrivateReferral$subgroup <- "PrivateReferral"

## PublicReferral only
l <- ROCPRC(PublicReferral)
PublicReferral <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
PublicReferral$subgroup <- "PublicReferral"

## DOTS only
l <- ROCPRC(DOTS)
DOTS <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
DOTS$subgroup <- "DOTS retested"

## Walk-in only
l <- ROCPRC(WalkIn)
WalkIn <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
WalkIn$subgroup <- "WalkIn"

## Community only
l <- ROCPRC(Community)
Community <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
Community$subgroup <- "Community"

## Contacts only
l <- ROCPRC(Contacts)
Contacts <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
Contacts$subgroup <- "Contacts"

## Female
l <- ROCPRC(Female)
Female <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
Female$subgroup <- "Female"

## Male
l <- ROCPRC(Male)
Male <- data.frame(matrix(unlist(l), nrow=5, byrow=T))
Male$subgroup <- "Male"
#### stratification by age AND prior TB history

aucTable <- rbind(all, New, Old, Young, Middle, Senior, PrivateReferral, PublicReferral, DOTS, WalkIn,Community, Contacts, Female, Male)

colnames(aucTable) <- c("AI.Algorithm", "AUC", "PRAUC", "Subgroup")
write.csv(aucTable, "Results/aucTable.csv")
rm(list=(ls()))
```

```{r ROC AUC with CI}
library(pROC)
ROC <- function(dataset){
  roc_CAD6 <- ci.auc(Xpert2Outcome_num ~ CAD4TB6, dataset)
  roc_qure <- ci.auc(Xpert2Outcome_num ~ qXRv3_100, dataset)
  roc_Lunit <- ci.auc(Xpert2Outcome_num ~ LunitScore, dataset)
  roc_JF1 <- ci.auc(Xpert2Outcome_num ~ JF1, dataset)
  roc_IF2 <- ci.auc(Xpert2Outcome_num ~ IF2, dataset)

  aucT <- as.data.frame(matrix(c(roc_CAD6[2], roc_CAD6[1], roc_CAD6[3], roc_qure[2], roc_qure[1], roc_qure[3], roc_Lunit[2], roc_Lunit[1], roc_Lunit[3], roc_JF1[2], roc_JF1[1], roc_JF1[3], roc_IF2[2], roc_IF2[1], roc_IF2[3]),  byrow=TRUE, ncol =  3))
  names(aucT) <- c("AUC", "AUCL", "AUCH")
  aucT$AI.Algorithm <- c("CAD4TB", "qXR", "Lunit INSIGHT CXR", "JF CXR-1", "InferReadDR")

  return(aucT)
}

## New only
New <- ROC(New)
New$subgroup <- "New cases"

## Old only
Old <- ROC(Old)
Old$subgroup <- "Previously treated cases"

## Young
Young <- ROC(Young)
Young$subgroup <- "Young age"

## Middle
Middle <- ROC(Middle)
Middle$subgroup <- "Middle age"

## Senior
Senior <- ROC(Senior)
Senior$subgroup <- "Old age"

## PrivateReferral only
PrivateReferral <- ROC(PrivateReferral)
PrivateReferral$subgroup <- "PrivateReferral"

## PublicReferral only
PublicReferral <- ROC(PublicReferral)
PublicReferral$subgroup <- "PublicReferral"

## DOTS only
DOTS <- ROC(DOTS)
DOTS$subgroup <- "DOTS retested"

## WalkIn only
WalkIn <- ROC(WalkIn)
WalkIn$subgroup <- "WalkIn"

## Community only
Community <- ROC(Community)
Community$subgroup <- "Community"

## Contacts only
Contacts <- ROC(Contacts)
Contacts$subgroup <- "Contacts"

## Female only
Female <- ROC(Female)
Female$subgroup <- "Female"

## Male only
Male <- ROC(Male)
Male$subgroup <- "Male"

ROCaucTable <- rbind(New, Old, Young, Middle, Senior, PrivateReferral, PublicReferral, DOTS, WalkIn,Community, Contacts, Female, Male)
View(ROCaucTable)
write.csv(ROCaucTable, "Results/ROCaucTable.csv", row.names = F)
```

```{r Bar chart}
rm(list = ls())

aucTable <- read.csv("Results/aucTable.csv")
aucTable <- aucTable[, -1]

aucTable$Subgroup <- factor(aucTable$Subgroup, levels = c("All","Young age", "Middle age", "Old age", "PrivateReferral", "PublicReferral", "DOTS retested", "WalkIn", "Community", "Contacts", "New cases", "Previously treated cases", "Female", "Male"))
### AUC
# age<-ggplot(aucTable[aucTable$Subgroup %in% c("Young age", "Middle age", "Old age"), ], aes(x=AI.Algorithm, y=AUC, fill=Subgroup)) +  geom_bar(stat="identity", width=0.6, position=position_dodge())+ geom_text(aes(label=AUC), position=position_dodge(width=0.9), vjust=0.05, hjust = 1.5, angle=90)+theme_minimal()+ theme(legend.position = "top")+  ylim(0, 0.95) + theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.95, 0.05), breaks = seq(0.65, 0.95, 0.05)) 
# 
# 
# referral<-ggplot(aucTable[aucTable$Subgroup %in% c("PrivateReferral", "PublicReferral", "DOTS retested", "WalkIn", "Community", "Contacts"), ], aes(x=AI.Algorithm, y=AUC, fill=Subgroup)) + geom_bar(stat="identity", width=0.7, position=position_dodge())+theme_minimal()+ geom_text(aes(label=AUC), position=position_dodge(width=0.9), vjust=0.05, hjust = 1.5, angle=90)+ theme(legend.position = "top")+  ylim(0, 0.95) + theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.95, 0.05), breaks = seq(0.65, 0.95, 0.05)) 
# 
# history<-ggplot(aucTable[aucTable$Subgroup %in% c("New cases", "Previously treated cases"), ], aes(x=AI.Algorithm, y=AUC, fill=Subgroup)) + geom_bar(stat="identity", width=0.5, position=position_dodge())+theme_minimal() + geom_text(aes(label=AUC), position=position_dodge(width=0.9), vjust=0.05, hjust = 1.5, angle=90) + theme(legend.position = "top") +  ylim(0, 0.95) + theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.95, 0.05), breaks = seq(0.65, 0.95, 0.05)) 
# 
# gender <- ggplot(aucTable[aucTable$Subgroup %in% c("Female", "Male"), ], aes(x=AI.Algorithm, y=AUC, fill=Subgroup)) + geom_bar(stat="identity", width=0.5, position=position_dodge())+ geom_text(aes(label=AUC), position=position_dodge(width=0.9), vjust=0.05, hjust = 1.5, angle=90) +theme_minimal() + theme(legend.position = "top") +  ylim(0, 0.95) + theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.95, 0.05), breaks = seq(0.65, 0.95, 0.05))

### AUC with CI
ROCaucTable <- read.csv("Results/ROCaucTable.csv")
ROCaucTable$AUC <- round(ROCaucTable$AUC, 4)

ROCaucTable$subgroup <- factor(ROCaucTable$subgroup, levels = c("All","Young age", "Middle age", "Old age", "PrivateReferral", "PublicReferral", "DOTS retested", "WalkIn", "Community", "Contacts", "New cases", "Previously treated cases", "Female", "Male"))

age <-ggplot(ROCaucTable[ROCaucTable$subgroup %in% c("Young age", "Middle age", "Old age"), ], aes(x=AI.Algorithm, y=AUC, fill=subgroup)) +  geom_bar(stat="identity", width=0.6, position=position_dodge()) + geom_errorbar(aes(ymin=AUCL, ymax=AUCH), width=.2, position=position_dodge(0.7)) + geom_text(aes(label=AUC), position=position_dodge(width=0.9), vjust=0.05, hjust = 1.5, angle=90) + theme_minimal()+ theme(legend.position = "top")+  ylim(0, 0.95) + theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.95, 0.05), breaks = seq(0.65, 0.95, 0.05)) 
age

referral<-ggplot(ROCaucTable[ROCaucTable$subgroup %in% c("PrivateReferral", "PublicReferral", "DOTS retested", "WalkIn", "Community", "Contacts"), ], aes(x=AI.Algorithm, y=AUC, fill=subgroup)) + geom_bar(stat="identity", width=0.7, position=position_dodge())+theme_minimal() + geom_text(aes(label=AUC), position=position_dodge(width=0.9), vjust=0.05, hjust = 1.5, angle=90)+ theme(legend.position = "top")+  ylim(0, 0.95) + theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.95, 0.05), breaks = seq(0.65, 0.95, 0.05)) + geom_errorbar(aes(ymin=AUCL, ymax=AUCH), width=.2, position=position_dodge(0.6)) 
referral

history<-ggplot(ROCaucTable[ROCaucTable$subgroup %in% c("New cases", "Previously treated cases"), ], aes(x=AI.Algorithm, y=AUC, fill=subgroup)) + geom_bar(stat="identity", width=0.5, position=position_dodge())+theme_minimal() + theme(legend.position = "top") +  ylim(0, 0.95) + geom_text(aes(label=AUC), position=position_dodge(width=0.9), vjust=0.05, hjust = 1.5, angle=90)+ theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.95, 0.05), breaks = seq(0.65, 0.95, 0.05)) + geom_errorbar(aes(ymin=AUCL, ymax=AUCH), width=.2, position=position_dodge(0.6))

gender<-ggplot(ROCaucTable[ROCaucTable$subgroup %in% c("Female", "Male"), ], aes(x=AI.Algorithm, y=AUC, fill=subgroup)) + geom_bar(stat="identity", width=0.5, position=position_dodge())+theme_minimal() + theme(legend.position = "top") +  ylim(0, 0.95) + geom_text(aes(label=AUC), position=position_dodge(width=0.9), vjust=0.05, hjust = 1.5, angle=90)+ theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.95, 0.05), breaks = seq(0.65, 0.95, 0.05)) + geom_errorbar(aes(ymin=AUCL, ymax=AUCH), width=.2, position=position_dodge(0.6))

### PRC
agePC<-ggplot(aucTable[aucTable$Subgroup %in% c("Young age", "Middle age", "Old age"), ], aes(x=AI.Algorithm, y=PRAUC, fill=Subgroup)) +  geom_bar(stat="identity", width=0.6, position=position_dodge())+ geom_text(aes(label=PRAUC), position=position_dodge(width=0.9), vjust=0.05, hjust = 1.5, angle=90)+theme_minimal()+ theme(legend.position = "top")+  ylim(0.3, 0.8) + theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.8, 0.05), breaks = seq(0.3, 0.8, 0.05))

referralPC<-ggplot(aucTable[aucTable$Subgroup %in% c("PrivateReferral", "PublicReferral", "DOTS retested", "WalkIn", "Community", "Contacts"), ], aes(x=AI.Algorithm, y=PRAUC, fill=Subgroup)) + geom_bar(stat="identity", width=0.7, position=position_dodge())+ geom_text(aes(label=PRAUC), position=position_dodge(width=0.9), vjust=0.05, hjust = 1.5, angle=90)+theme_minimal()+ theme(legend.position = "top")+  ylim(0.3, 0.8) + theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.8, 0.05), breaks = seq(0.3, 0.8, 0.05))

historyPC<-ggplot(aucTable[aucTable$Subgroup %in% c("New cases", "Previously treated cases"), ], aes(x=AI.Algorithm, y=PRAUC, fill=Subgroup)) + geom_bar(stat="identity", width=0.5, position=position_dodge())+ geom_text(aes(label=PRAUC), position=position_dodge(width=0.9), vjust=0.05, hjust = 1.5, angle=90)+theme_minimal() + theme(legend.position = "top") +  ylim(0.3, 0.8) + theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.8, 0.05), breaks = seq(0.3, 0.8, 0.05))

genderPC <- ggplot(aucTable[aucTable$Subgroup %in% c("Female", "Male"), ], aes(x=AI.Algorithm, y=PRAUC, fill=Subgroup)) + geom_bar(stat="identity", width=0.5, position=position_dodge())+ geom_text(aes(label=PRAUC), position=position_dodge(width=0.9), vjust=0.05, hjust = 1.5, angle=90)+theme_minimal() + theme(legend.position = "top") +  ylim(0, 0.95) + theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.95, 0.05), breaks = seq(0.65, 0.95, 0.05))

# 
# tiff("Results/PRC ROC AUC plot 3.tif", width = 20, height = 8, units = "in", res = 110)
# require(gridExtra)
# grid.arrange(age, referral, history, gender, agePC, referralPC, historyPC, genderPC, ncol=4)
# dev.off()

tiff("Results/Figure3_250.tif", width = 12, height = 14, units = "in", res = 250)
require(gridExtra)
grid.arrange(age, referral, history, gender, agePC, referralPC, historyPC, genderPC, ncol=2)
dev.off()

# tiff("Results/PRC AUC plot.tif", width = 9, height = 8, units = "in", res = 100)
# grid.arrange(agePC, referralPC, historyPC, genderPC, ncol=2)
# dev.off()
```

```{r}
ROCaucTable <- read.csv("Results/ROCaucTable.csv")
ROCaucTable$AUC <- round(ROCaucTable$AUC, 3)

ROCaucTable$subgroup <- factor(ROCaucTable$subgroup, levels = c("All","Young age", "Middle age", "Old age", "PrivateReferral", "PublicReferral", "DOTS retested", "WalkIn", "Community", "Contacts", "New cases", "Previously treated cases", "Female", "Male"))

### AUC with CI
age <-ggplot(ROCaucTable[ROCaucTable$subgroup %in% c("Young age", "Middle age", "Old age"), ], aes(x=AI.Algorithm, y=AUC, fill=subgroup)) +  geom_bar(stat="identity", width=0.6, position=position_dodge()) + geom_errorbar(aes(ymin=AUCL, ymax=AUCH), width=.2, position=position_dodge(0.7)) + geom_text(aes(label=AUC), position=position_dodge(width=0.9), vjust=0.05, hjust = 1.5, angle=90) + theme_minimal()+ theme(legend.position = "top")+  ylim(0, 0.95) + theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.95, 0.05), breaks = seq(0.65, 0.95, 0.05)) 
age

referral<-ggplot(ROCaucTable[ROCaucTable$subgroup %in% c("PrivateReferral", "PublicReferral", "DOTS retested", "WalkIn", "Community", "Contacts"), ], aes(x=AI.Algorithm, y=AUC, fill=subgroup)) + geom_bar(stat="identity", width=0.7, position=position_dodge())+theme_minimal() + geom_text(aes(label=AUC), position=position_dodge(width=0.9), vjust=0.05, hjust = 1.5, angle=90)+ theme(legend.position = "top")+  ylim(0, 0.95) + theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.95, 0.05), breaks = seq(0.65, 0.95, 0.05)) + geom_errorbar(aes(ymin=AUCL, ymax=AUCH), width=.2, position=position_dodge(0.6)) 
referral

history<-ggplot(ROCaucTable[ROCaucTable$subgroup %in% c("New cases", "Previously treated cases"), ], aes(x=AI.Algorithm, y=AUC, fill=subgroup)) + geom_bar(stat="identity", width=0.5, position=position_dodge())+theme_minimal() + theme(legend.position = "top") +  ylim(0, 0.95) + geom_text(aes(label=AUC), position=position_dodge(width=0.9), vjust=0.05, hjust = 1.5, angle=90)+ theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.95, 0.05), breaks = seq(0.65, 0.95, 0.05)) + geom_errorbar(aes(ymin=AUCL, ymax=AUCH), width=.2, position=position_dodge(0.6))

gender<-ggplot(ROCaucTable[ROCaucTable$subgroup %in% c("Female", "Male"), ], aes(x=AI.Algorithm, y=AUC, fill=subgroup)) + geom_bar(stat="identity", width=0.5, position=position_dodge())+theme_minimal() + theme(legend.position = "top") +  ylim(0, 0.95) + geom_text(aes(label=AUC), position=position_dodge(width=0.9), vjust=0.05, hjust = 1.5, angle=90)+ theme(panel.grid.minor = element_line(size=0.5)) + scale_y_continuous(minor_breaks = seq(0 , 0.95, 0.05), breaks = seq(0.65, 0.95, 0.05)) + geom_errorbar(aes(ymin=AUCL, ymax=AUCH), width=.2, position=position_dodge(0.6))

tiff("Results/Figure4.tif", width = 12, height = 7, units = "in", res = 250)
require(gridExtra)
grid.arrange(age, referral, history, gender, ncol=2)
dev.off()

```